Convergence rate of MAP estimates \\ for the exponential family

### 2020-11-24 main.tex
Comment on the prior work with relative smoothness

\usepackage[round]{natbib}
\renewcommand{\bibname}{References}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

@@

\section{Stochastic Mirror Descent}
It turns out that the MAP can also be seen as the trajectory of stochastic mirror descent
\begin{align}
    \hat \natp_{n+1} 
    &= \argmin_\natp 
    - \langle T(x_{t+1}), \natp \rangle  + \logpartition(\natp) + (n_0 + n) \bregman(\natp || \natp_n) \\
    %&(n_0+n)\left(\logpartition(\natp) - \logpartition(\natp_n) - \langle \nabla \logpartition(\natp_n)  , \natp - \natp_n \rangle \right)  \\
    &= \argmin_\natp - \langle T(x_{t+1}), \natp \rangle   + \logpartition(\natp_n) + \langle \nabla \logpartition(\natp_n)  , \natp - \natp_n \rangle + (n_0+n + 1) \bregman(\natp || \natp_n) \\
    &= \argmin_\natp \linear_f(\natp;\natp_n) + (n_0 + n + 1) \bregman(\natp || \natp_n)
\end{align}
where $\linear_f(\natp;\natp_n)$ is the linearization of $f$ at $\natp_n$ evaluated at $\natp$. 
This is the formula for stochastic mirror descent (SMD) applied to $f$ with mirror map $\logpartition$ and step-size $\gamma_n = \inv{n_0 + n + 1}$.

In the classic setting, SMD is studied under strong-convexity assumption on the mirror map $\logpartition$ \RLP{cite {bubeck}}.
In our setting this is not always true -- eg gaussians.
However a recent and fast-expanding body of work is concerned with a new assumption: relative smoothness and relative strong-convexity. 
For mirror descent, there are well linear and sublinear rate, even when the function is not smooth or strongly convex ! Beautiful. Cite nesterov, and teboulle, and the first birnbaum.

For stochastic mirror descent, there is  a sublinear rate (cite richtarik), but under a weird assumption on gradient's variance. Still need to check “Relative Continuity” for Non-Lipschitz Nonsmooth Convex Optimization Using Stochastic (or Deterministic) Mirror Descent".

@@

\bibsection
\bibliography{}

>>>>

\usepackage[round]{natbib}
\renewcommand{\bibname}{References}
\renewcommand{\bibsection}{\subsection*{\bibname}}

@@

\section{Stochastic Mirror Descent}
It turns out that the MAP can also be seen as the trajectory of stochastic mirror descent
\begin{align}
    \hat \natp_{n+1} 
    &= \argmin_\natp 
    - \langle T(x_{t+1}), \natp \rangle  + \logpartition(\natp) + (n_0 + n) \bregman(\natp || \natp_n) \\
    %&(n_0+n)\left(\logpartition(\natp) - \logpartition(\natp_n) - \langle \nabla \logpartition(\natp_n)  , \natp - \natp_n \rangle \right)  \\
    &= \argmin_\natp - \langle T(x_{t+1}), \natp \rangle   + \logpartition(\natp_n) + \langle \nabla \logpartition(\natp_n)  , \natp - \natp_n \rangle + (n_0+n + 1) \bregman(\natp || \natp_n) \\
    &= \argmin_\natp \linear_f(\natp;\natp_n, x_{t+1}) + (n_0 + n + 1) \bregman(\natp || \natp_n)
\end{align}
where $\linear_f(\natp;\natp_n, x_{t+1})$ is the linearization of $f$ at $\natp_n$ evaluated at $\natp$ with randomness coming from $x_{t+1}$. 
This is the formula for stochastic mirror descent (SMD) applied to $f$ with mirror map $\logpartition$ and step-size $\gamma_n = \inv{n_0 + n + 1}$.

In the classic setting, SMD is studied under strong-convexity assumption on the mirror map $\logpartition$ \citep{bubeck2015convex}.
In our setting this is not always true -- eg gaussians.
However a recent and fast-expanding body of work is concerned with a new assumption: relative smoothness and relative strong-convexity. $f$ is $\mu$ strongly-convex and $L$-smooth relative to $h$ if
\begin{align}
	f(x) + \langle \nabla f(x) , y-x \rangle + \mu \cB_h(y||x)
	\leq f(y) \\
	\text{and} \quad 
	f(y) \leq 	f(x) + \langle \nabla f(x) , y-x \rangle + L\cB_h(y||x), \forall x,y \\
	\iff \mu\cB_h(y||x) \leq \cB_f(y||x) \leq L \cB_h(y||x), \forall x,y \\
	\iff \mu \nabla^2 h(x) \leq \nabla^2 f(x) \leq L \nabla^2 h(x),  \forall x
\end{align} 
where the last equivalence holds only when $f$ and $h$ are twice differentiable. 
This sweet generalization of smoothness and strong-convexity transfers the Loewner partial order between matrices to functions, via the Hessian. 
As such it can be applied to many functions that were out of reach for $\ell^2$ norm, by taking the appropriate reference function.  For instance $h(x) = -\log(x)$ or $h(x) = x^4$. 
As early as 2011, \citet{birnbaum2011distributed} showed $O(\inv{t})$ convergence rate for mirror descent under smoothness assumption relative to the mirror map. 
More precisely, he proved the simple finite rate
\begin{align}
	f(x_t) - f(x_*) \leq \frac{L \cB_h(x_* || x_0) }{t} \; .
\end{align}
These notions were rediscovered and expanded by \citet{bauschke2017descent} and \citet{lu2018relatively}. If you need to read one, pick \citet{lu2018relatively} -- I found it much much easier and more enjoyable to read.  This latter paper also derived a linear convergence rate for mirror descent under relative smoothness and strong-convexity, with the relative condition number $\frac{L}{\mu}$ appearing.

Now our setting is Stochastic Mirror Descent (SMD), meaning at each step we observe a random unbiased estimate gradient. This setting was studied by \citet{hanzely2018fastest}, who proved in the smooth strongly-convex case with tail averaging :  with constant step-size, linear convergence down to a variance ball, and with step-size $\gamma_t = n_0 + t$ a rate $\tilde O(\inv{t})$. These results match the rates for standard SGD.
This is very interesting for us, but the variance hyper-parameter is oddly defined. Let $g_t$ be the random gradient at step $t$  (coming from data point $x_t$), and $\natp_{t+1}$ the next iterate. Then the variance bound $\sigma^2$ is a lower bound on the covariance between $g_t$ and $\natp_{t+1}$. 
\begin{align}
	\Cov(g_t,\natp_{t+1}) 
	&= \expect{\langle g_t, \natp_{t+1} \rangle}  - \langle \E[g_t], \E[\natp_{t+1}] \rangle \geq -  \gamma_t \sigma^2, \forall t \\
	 &= \expect{\langle g_t - \nabla f(\natp_t), \natp_{t+1}  - \natp \rangle}, (\forall \natp)
\end{align}
where $\gamma_t$ is the step-size and expectations are conditional on $\natp_t$.
Still need to check €œRelative Continuity€ for Non-Lipschitz Nonsmooth Convex Optimization Using Stochastic (or Deterministic) Mirror Descent, which makes different hypothesis and is well-written.

@@

\bibsection
\bibliographystyle{apalike}
\bibliography{references.bib}
